---
title: "Bacteria and DOM experiments"
author: "Nick Baetge"
date: "compiled most recently on `r format(Sys.time(), '%d %B, %Y')`"
output: github_document

knit: (function(inputFile, encoding) {
      out_dir <- "../knitted";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
---

```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(hms)
library(lubridate)
library(purrr)
library(patchwork)
library(janitor)
library(gt)
```

```{r define path to save data}
prod_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/p5_bacteria_dom_exps.csv"
bcodmo_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/bco_dmo/BACTERIA_DOM.csv"
```

# Bacterial cell data

## Import and tidy data

```{r define paths to fcm data}
fcm_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/p4_fcm_plate_data"
filter_vols <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_filter_vols.xlsx"
time_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_bact_dom_sampletimes.csv"
```

```{r import and tidy sample time data, message = F}
dt <- read_csv(time_path) %>%
  mutate_at(vars(datetime_utc), ymd_hms) %>%
  rename(dt = 3) %>%
  group_by(exp) %>%
  mutate(
    interv = interval(first(dt), dt),
    dur = as.duration(interv),
    days = round(time_length(dur, unit = "day"), 2)
  ) %>%
  select(exp, tp, dt, days)
```

import data and clean based on run notes:

- B2 run 1 (and dil): timepoint 1 and 2 are switched
- B3_14 run 1 dil: B14-2-4-0 didn't add sample so data is in well H12
- B16 run 2 dil:  B16-V2-4-4 didn't add sample so data is in well G1
- B3_14 run 2 B14_V1_3 was a 2 ml sample instead of a 1 ml sample, also extra 1.2 vial (could be 1.2 GF75)
- Filtrates run 2: missing vial for B3 1.2 GF75 

```{r import and tidy fcm data, message = F}
fcm <-
  list.files(path = fcm_path,
             pattern = "*.csv",
             full.names = T) %>%
  map_df( ~ read_csv(., col_types = cols(.default = "c"))) %>%
  mutate_at(vars(tp, dil, cells_ml), as.numeric) %>%
  mutate(tp = ifelse(
    name %in% c(
      "B2_dil-1.fcs",
      "B2-7.fcs",
      "B2_dil-2.fcs",
      "B2-8.fcs",
      "B2_dil-3.fcs",
      "B2-9.fcs",
      "B2_dil-4.fcs",
      "B2-10.fcs"
    ),
    2,
    tp
  ),
  tp = ifelse(
    name %in% c("B2-11.fcs", "B2-12.fcs", "B2-13.fcs", "B2-14.fcs"),
    1,
    tp
  )) %>%   # sample switch
  mutate(btl = ifelse(name == "B3_14-1.fcs", "1.2GF75", btl)) %>%    # sample switch
  filter(!name %in% c(
    "B3_14_dil-33.fcs",
    "B15-2.fcs",
    "B16_r2-25.fcs",
    "B16_r2-25.fcs"
  )) %>%    # samples had 0 events
  mutate(
    trt = case_when(
      btl %in% c("1", "V1", "1.2T", "1.2DNA", "1.2GF75", "1DNA", "1GF75") ~ 1,
      btl %in% c("2", "V2", "2DNA", "2GF75") ~ 2,
      btl %in% c("3", "V3", "3DNA", "3GF75") ~ 3,
      btl %in% c("4", "V4", "4DNA", "4GF75") ~ 4
    ),
    type = ifelse(btl %in% c("1", "2", "3", "4"), "5L Bottle", "40 mL Vial"),
    desc = case_when(
      trt == 1 ~ "Control",
      trt == 2 ~ "Thomas Fire Ash",
      trt == 3 ~ "Low Temp. Ash",
      trt == 4 ~ "High Temp. Ash"
    ),
    .after = btl
  ) %>%
  arrange(exp, trt, tp) 
```

## Subtract sybr stain

```{r calculate sybr correction}
sybr <- fcm %>%
  filter(exp == "TE+SYBR") %>%
  mutate(med_sybr = round(median(cells_ml))) %>%
  select(med_sybr) %>%
  distinct() %>%
  as.numeric()
```

```{r correct for sybr}
fcm_corr <- fcm %>%
  select(name, exp, btl, trt, desc, type, tp, dil, cells_ml) %>%
  arrange(exp, btl, tp) %>%
  filter(!exp %in% c("TE", "TE+SYBR")) %>%
  mutate(cells_ml = cells_ml - sybr,
         cells_ml = ifelse(cells_ml < 0, 0, cells_ml)) 
```

## Assess vial v bottle cell dynamics

```{r regression of vial cells and bottle cells, message = F}
reg_data <- fcm %>%
  filter(btl %in% c("1", "2", "3", "4", "V1", "V2", "V3", "V4")) %>%
  group_by(exp, trt, btl, type, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml))) %>%
  ungroup() %>%
  select(exp, trt, desc, type, tp, dil, mean_cells_ml) %>%
  distinct() %>%
  pivot_wider(
    id_cols = c(exp, trt, desc, tp, dil),
    names_from = type,
    values_from = mean_cells_ml
  ) %>%
  drop_na(7)

reg <-
  lmodel2::lmodel2(reg_data$`40 mL Vial` ~ reg_data$`5L Bottle`, nperm = 99)

reg
```
The cell dynamics occurring in the bottle and the vials track well with each other. we can use averages between the two for subsequent measurements. so cells per ml will reflect the average of two subsamples from the vials and two subsamples from the bottle (n = 4)

## Extract experimental data

```{r extract experimental data, message = F}
# apply cell counts from bottles at T0 as vial values for T0
fcm_40mlt0 <- fcm %>%
  filter(btl %in% c("1", "2", "3", "4")) %>%
  group_by(exp, trt, btl, type, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml)),
         sd_cells_ml = round(sd(cells_ml))) %>%
  ungroup() %>%
  filter(tp == 0) %>%
  mutate(type = "40 mL Vial",
         btl = paste("V", btl, sep = ""))

#extract experimental data and merge with datetime data
fcm_exp <- fcm %>%
  filter(btl %in% c("1", "2", "3", "4", "V1", "V2", "V3", "V4")) %>%
  group_by(exp, trt, btl, type, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml)),
         sd_cells_ml = round(sd(cells_ml))) %>%
  ungroup() %>%
  bind_rows(., fcm_40mlt0) %>%
  select(-name) %>%
  arrange(exp, btl, tp) %>%
  left_join(., dt) %>%
  select(exp:tp, dt, days, everything()) %>%
  mutate(hours = days * 24, .after = days)

#calculate mean data
fcm_exp_means <- fcm_exp %>%
  select(-c(type, dil, btl, mean_cells_ml, sd_cells_ml)) %>%
  arrange(exp, trt, tp) %>%
  group_by(exp, trt, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml)),
         sd_cells_ml = round(sd(cells_ml))) %>%
  select(-cells_ml) %>%
  distinct() %>% 
  ungroup() 
```

## Estimate growth rates (d-1) in exponential phase

```{r estimate growth rates, message = F}
fcm_gr <- fcm_exp %>%
  filter(
    exp == "B1" & tp <= 3 |
      exp == "B2" & tp <= 2 |
      exp == "B3" & tp <= 2 |
      exp == "B14" & tp <= 2 |
      exp == "B15" &  tp %in% c(1, 2, 3) |
      exp == "B16" & tp %in% c(1, 2, 3)
  ) %>%
  group_by(exp, btl, trt) %>%
  mutate(gr_d = (log(last(mean_cells_ml)) - log(first(mean_cells_ml))) / (last(days) - first(days))) %>%
  select(exp:desc, gr_d) %>%
  distinct() %>%
  group_by(exp, trt, desc) %>%
  summarise_at(vars(gr_d), list(mean = mean, sd = sd)) %>%
  ungroup() %>%
  rename(mean_gr_d = mean,
         sd_gr_d = sd)

p5_fcm <- fcm_gr  %>%
  left_join(fcm_exp_means, .) %>%
  select(exp,
         desc,
         tp,
         dt,
         days,
         hours,
         mean_cells_ml,
         sd_cells_ml,
         mean_gr_d,
         sd_gr_d) %>%
  distinct()
```

## Estimate filter retention for bacterial organic carbon and dna samples

```{r extract filtration data and calculate retention, message = F}
fcm_gf75_dna <- fcm %>%
  filter(tp %in% c(0, 2, 5)) %>%
  mutate(cells_ml = ifelse(
    name %in% c(
      "Filtrates-21.fcs",
      "Filtrates_r2-21.fcs",
      "Filtrates_r2-19.fcs",
      "Filtrates-19.fcs"
    ),
    NA,
    cells_ml
  )) %>%  # suspect samples. these gf75 filtrates had more cells than any other sample in the dataset.
  arrange(exp, tp, btl) %>%
  group_by(exp, btl, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml))) %>%
  ungroup() %>%
  select(-c(name, dil, cells_ml)) %>%
  distinct() %>%
  filter(!btl %in% c(1, 2, 3, 4) | !tp == 0) %>%
  group_by(exp, trt, tp) %>%
  mutate(
    total = ifelse(
      str_detect(btl, "DNA") |
        str_detect(btl, "GF75"),
      NA,
      mean_cells_ml
    ),
    gf75_filtrate = ifelse(str_detect(btl, "GF75"), mean_cells_ml, NA),
    dna_filtrate = ifelse(str_detect(btl, "DNA"), mean_cells_ml, NA)
  ) %>%
  fill(c(total:dna_filtrate), .direction = "updown") %>%
  ungroup() %>%
  select(exp, trt, desc, tp, total:dna_filtrate) %>%
  distinct() %>%
  group_by(exp, trt, tp) %>%
  mutate(mean_total = round(mean(total, na.rm = T)), .after = tp) %>%
  select(-total) %>%
  distinct() %>%
  rename(total = mean_total) %>%
  left_join(readxl::read_xlsx(filter_vols, sheet = 1), .) %>%
  mutate(
    gf75_total = total * gf75_vol,
    dna_total = total * dna_vol,
    .before = gf75_filtrate
  ) %>% #total cells in 1L
  mutate_at(vars(dna_filtrate), ~ . * dna_vol) %>%
  mutate_at(vars(gf75_filtrate), ~ . * gf75_vol) %>%
  select(-total) %>%
  mutate(
    gf75_filter = gf75_total - gf75_filtrate,
    dna_filter = dna_total - dna_filtrate,
    gf75_percent_retention = round(gf75_filter / gf75_total * 100),
    dna_percent_retention = round(dna_filter / dna_total * 100)
  )
```

```{r filter retention of cells summary statistics}
summary(fcm_gf75_dna)
sd(fcm_gf75_dna$gf75_percent_retention, na.rm = T)
sd(fcm_gf75_dna$dna_percent_retention)
```

# Bacterial organic carbon data

## Import and tidy data

```{r define paths to boc data}
chn_path <- "/Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_chn.csv"
```

```{r import data, message = F, warning = F}
p5_boc <- read_csv(chn_path) %>% 
  left_join(., readxl::read_xlsx(filter_vols, sheet = 2)) %>% 
  select(1, 2, 4, 5, 7) %>% 
  rename(rundate = 1,
         exp_sample_tp_rep = 2, 
         c_ug = 3, 
         n_ug = 4) %>% 
  separate(., 2, into = c("exp", "sample", "tp", "rep"), sep = "-") %>% 
  separate(., 3, into = c("sample", "filter"), sep = -1) %>% 
  group_by(exp, sample, tp, rep) %>% 
   mutate(raw_c = round(sum(c_ug),1),
         raw_n = round(sum(n_ug),1)) %>% 
  select(-c(filter, c_ug, n_ug)) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(c_blank = round(mean(raw_c[exp == "BLK"], na.rm = T), 1),
         n_blank = round(mean(raw_n[exp == "BLK"], na.rm = T), 1),
         c_blank_sd = round(sd(raw_c[exp == "BLK"], na.rm = T), 1),
         n_blank_sd = round(sd(raw_n[exp == "BLK"], na.rm = T), 1)) %>% 
  filter(!exp == "BLK") %>% 
  mutate(c_ug = raw_c - c_blank,
         n_ug = raw_n - n_blank) %>% 
  filter(!c_ug < 0) %>% 
  mutate(c_mg = c_ug / 10^3,
         n_mg = n_ug / 10^3,
         c_umol = c_mg * 10^3 / 12,
         n_umol = n_mg * 10^3 / 14,
         vol_m3 = vol / 10^6,
         c_mg_m3 = round(c_mg / vol_m3, 1),
         n_mg_m3 = round(n_mg / vol_m3, 1),
         c_umol_l = c_mg_m3/12000,
         n_umol_l = n_mg_m3/14000,
         cn =  round(c_umol_l/n_umol_l, 1)) %>% 
  select(exp, sample, tp, c_mg:cn) %>% 
  rename(trt = sample) %>% 
  group_by(exp, trt, tp) %>% 
  summarize(across(
    .cols = c_mg:cn,
    .fns = list(mean = mean, sd = sd),
    na.rm = TRUE,
    .names = "{fn}_{col}"
  )) %>% 
  ungroup() %>% 
  mutate_at(vars(tp), as.numeric) %>% 
  mutate(trt = ifelse(trt == 1.2, 1, trt)) %>% 
  group_by(exp, trt) %>% 
  group_modify(~ add_row(., tp = 0 )) %>% 
  arrange(exp,  tp) %>% 
  group_by(exp, tp) %>% 
  fill(c(3:23), .direction = "down") %>% 
  ungroup() %>% 
  distinct() %>% 
  arrange(exp, trt, tp) %>% 
  mutate_at(vars(trt), as.numeric) %>% 
  left_join(., fcm_gf75_dna %>% select(exp, trt, desc, tp, gf75_filter)) %>% 
  group_by(exp, tp) %>% 
  fill(gf75_filter, .direction = "downup") %>% 
  ungroup() %>% 
  rename(gf75_cells = gf75_filter) %>% 
  select(exp, trt, desc, tp, gf75_cells, everything()) %>% 
  mutate(fgC_cell = (mean_c_umol * 12 * 10^9)/gf75_cells,
         fgN_cell = (mean_n_umol * 14 * 10^9)/gf75_cells ) %>% 
  ungroup() %>% 
  mutate(desc = case_when(trt == 1 ~ "Control",
                          trt == 2 ~ "Thomas Fire Ash",
                          trt == 3 ~ "Low Temp. Ash",
                          trt == 4 ~ "High Temp. Ash")) %>% 
  select(exp, trt, desc, tp, fgC_cell, fgN_cell) %>% 
  left_join(p5_fcm,.) %>%
  select(-trt)  %>% 
  rename(trt = desc)  %>% 
  group_by(exp, trt) %>%
  mutate(fgC_cell = ifelse(exp == "B2" & tp == 3, fgC_cell[tp == 2], fgC_cell)) %>% 
  mutate(interp_fgC_cell = zoo::na.approx(fgC_cell, dt, na.rm = F)) %>% 
  mutate(boc = (mean_cells_ml * interp_fgC_cell) / (12 * 10^6)) %>% 
  ungroup() %>% 
  mutate(interp_bocdoc = ifelse(is.na(fgC_cell) | exp == "B2" & tp == 3, T, F))
```

c_blank = 15.3 +/- 5.1, n = 9
n_blank = 7.1 +/- 1, n = 9 

# DOC data

## Import and tidy data

```{r define paths to doc data}
doc_path <- "//Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_doc.xlsx"
```


```{r tidy and derive bge and doc bioav, message = F}
doc <- readxl::read_xlsx(doc_path, skip = 27) %>%
  drop_na() %>%
  select(1, 3) %>%
  rename(id = 1,
         toc = 2) %>%
  separate(id, into = c("exp", "btl", "tp"), sep = "-") %>%
  group_by(exp, btl, tp) %>%
  mutate(mean_toc = mean(toc),
         sd_toc = sd(toc),
         tp = as.numeric(tp)) %>%
  ungroup() %>%
  select(-toc) %>%
  distinct() %>%
  mutate(
    trt =  case_when(
      btl == 1 ~ "Control",
      btl == 2 ~ "Thomas Fire Ash",
      btl == 3 ~ "Low Temp. Ash",
      btl == 4 ~ "High Temp. Ash"
    ),
    .after = btl
  ) %>%
  select(-btl) %>%
  left_join(p5_boc, .) %>%
  mutate(doc = mean_toc - boc) %>% 
  mutate(toc_unc = sd_toc / sqrt(3),
         mean_toc_unc = round(mean(toc_unc), 1)) 


#inclusion points

incl <- doc %>% 
  select(exp, trt, tp, days) %>%
  mutate(bge_inclusion = ifelse(exp == "B16" & trt == "Control" & tp %in% c(0, 1, 2, 3, 4, 5) |
                                  exp == "B16" & trt != "Control" & tp %in% c(0, 1, 2, 3, 4, 5) |
                                  exp == "B15" & tp %in% c(0, 1, 2, 3) |
                                  exp == "B1"  & tp %in% c(0, 1, 2, 3, 4, 5) |
                                  exp == "B14" & trt != "Control" & tp %in% c(0, 1, 2, 3, 4) |
                                  exp == "B14" & trt == "Control" & tp %in% c(0, 1, 2, 3, 4, 5) |
                                  exp == "B3"  & tp %in% c(0, 1, 2) |
                                  exp == "B2" & tp %in% c(0, 1, 2, 3),
                                T, F)) %>% 
  filter(bge_inclusion == T)


#calculate boc rate of change as slope
boc_rate <- doc %>% 
  select(exp, trt, tp, days, boc) %>%
  drop_na(boc) %>%
  left_join(., incl) %>% 
  drop_na(bge_inclusion) %>% 
  group_by(exp, trt) %>%
  mutate(group_id = cur_group_id()) %>%
  ungroup() %>%
  group_split(group_id) %>%
  map( ~ lm(boc ~ days, data = .x)) %>%
  map_df(broom::tidy, .id = "group_id") %>%
  filter(term == "days") %>%
  select(group_id, estimate) %>%
  rename(boc_rate = estimate) %>%
  mutate_at(vars(group_id), as.numeric) %>%
  ungroup() %>% 
  left_join(
    .,
    doc %>%
      select(exp, trt) %>% distinct() %>%
      group_by(exp, trt) %>%
      mutate(group_id = cur_group_id()) %>%
      ungroup()
  ) %>% 
  select(-group_id)

#calculate doc rate of change as slope
doc_rate <- doc %>%
  select(exp, trt, tp, days, doc) %>%
  drop_na(doc) %>%
  left_join(., incl) %>% 
  drop_na(bge_inclusion) %>% 
  group_by(exp, trt) %>%
  mutate(group_id = cur_group_id()) %>%
  ungroup() %>%
  group_split(group_id) %>%
  map( ~ lm(doc ~ days, data = .x)) %>%
  map_df(broom::tidy, .id = "group_id") %>%
  filter(term == "days") %>%
  select(group_id, estimate) %>%
  rename(doc_rate = estimate) %>%
  mutate_at(vars(group_id), as.numeric) %>%
  mutate_at(vars(doc_rate), abs) %>%
  ungroup() %>% 
  left_join(
    .,
    doc %>%
      select(exp, trt) %>% distinct() %>%
      group_by(exp, trt) %>%
      mutate(group_id = cur_group_id()) %>%
      ungroup()
  ) %>% 
  select(-group_id)

#calculate bges and doc bioavailability
p5_bge_bioav <- doc %>% 
  left_join(., boc_rate) %>% 
  left_join(., doc_rate) %>% 
  group_by(exp, trt) %>% 
  mutate(doc_rem = ifelse(is.na(doc[tp == 5]), first(doc) - doc[tp == 2], first(doc) - last(doc) ),
         doc_bdl = ifelse(doc_rem <= 0.8, T, F),
         bge = ifelse(doc_bdl == F, boc_rate/doc_rate, NA)) %>% 
  ungroup() %>% 
  group_by(exp, tp) %>% 
  mutate(control_doc = doc[trt == "Control"]) %>% 
  ungroup() %>% 
  mutate(amend_doc = doc - control_doc) %>% 
  group_by(exp, trt) %>% 
  mutate(amend_doc_bioav = ifelse(trt != "Control", (amend_doc[tp == 0] - last(amend_doc, na_rm = T)) / amend_doc[tp == 0]* 100, NA )) %>% 
  ungroup() %>% 
  mutate(exp = paste("P", exp, sep = ""))
```

# Merge data with contextual in situ data 

```{r define paths to context data, message = F}
context_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/p1_insitu_index.csv"

p5_prod <- read_csv(context_path) %>% 
  select(stn, exp, lat, lon,  cluster, composite_z) %>% 
  left_join(., p5_bge_bioav) %>% 
  arrange(cluster) %>% 
  drop_na(trt) %>% 
  left_join(., incl %>% mutate(exp = gsub("B", "PB", exp)))
```

```{r}
bco_dmo <- p5_prod %>% 
  select(stn, lat, lon, dt, trt, mean_cells_ml, sd_cells_ml, boc, mean_toc, sd_toc) %>% 
  rename(bact_cells = mean_cells_ml,
         sd_bact_cells = sd_cells_ml) %>% 
  arrange(stn, trt, dt)
```


```{r filter data needed to run some stats}
data4stats <- p5_prod %>% 
  select(exp, trt, cluster, mean_gr_d, bge, doc_rate, amend_doc, amend_doc_bioav) %>%
  distinct() %>% 
  # mutate(bge = ifelse(is.na(bge), mean(bge, na.rm = T), bge)) %>% 
  rename(bact_mu = 4) %>% 
  mutate(trt2 = ifelse(trt == "Control", "Control", "Ash leachate")) 
```

```{r pivot data for plotting curves, message = F}
data4curves <- p5_prod %>%
  select(exp, trt, cluster, days, bge_inclusion, mean_toc, boc) %>%
  rename(toc = mean_toc) %>%
  pivot_longer(toc:boc, names_to = "var", values_to = "val") %>%
  mutate(
    var = case_when(
      var == "toc" ~ "bold(TOC~(µmol~C~L^-1))",
      var == "boc" ~ "bold(BOC~(µmol~C~L^-1))"
    )
  ) %>%
  left_join(
    .,
    p5_prod %>%
      select(exp, trt, days, bge_inclusion,  cluster, sd_toc) %>%
      mutate(var = "bold(TOC~(µmol~C~L^-1))") %>%
      rename(sd = sd_toc)
  ) %>%
   bind_rows(
    .,
    p5_prod %>%
      select(exp, trt, days, bge_inclusion,  cluster, doc, sd_toc) %>%
      mutate(var = "bold(DOC~(µmol~C~L^-1))") %>%
      rename(val = doc,
             sd = sd_toc)
  ) %>%
  bind_rows(
    .,
    p5_prod %>%
      select(exp, trt, days,  bge_inclusion, cluster, mean_cells_ml, sd_cells_ml) %>%
      mutate_at(vars(mean_cells_ml, sd_cells_ml), ~ . * 10 ^ 3) %>%
      mutate(var = "bold(Cells~(L^-1))") %>%
      rename(val = mean_cells_ml,
             sd = sd_cells_ml)
  ) %>%
  arrange(exp, trt, days, var) %>%
  mutate(bge_inclusion = ifelse(is.na(bge_inclusion), F, bge_inclusion)) %>% 
  mutate(bge_inclusion = ifelse(var %in% c("bold(TOC~(µmol~C~L^-1))", "bold(Cells~(L^-1))"), F, bge_inclusion)) %>% 
  mutate(cluster = case_when(cluster == "Low biomass" ~ "Low~biomass",
                             cluster == "High biomass" ~ "High~biomass")) 
```

# Save data

```{r save data}
write_csv(p5_prod, prod_path)
write_csv(bco_dmo, bcodmo_path)
```

# Plot data

```{r custom theme for plots}
custom.theme <- theme(
  legend.position = "top",
  legend.title = element_text(size = 35),
  legend.key.size = unit(3, "cm"),
  legend.key.spacing.x = unit(0.75, "cm"),
  legend.text = element_text(size = 38),
  legend.box = "vertical",
  legend.margin = margin(),
  axis.title = element_text(size = 44, face = "bold"),
  axis.text = element_text(size = 38),
  panel.spacing.x = unit(0.5, "cm"),
  panel.spacing.y = unit(1, "cm"),
  strip.text.x = element_text(
    size = 42,
    color = "white",
    face = "bold"
  ),
  strip.text.y = element_text(
    size = 35,
    color = "white",
    face = "bold"
  ),
  strip.background = element_rect(
    color = "black",
    fill = alpha('black', 0.7),
    linewidth = 1.5,
    linetype = "solid"
  )
)

pal4 = c("#785EF0", "#DC267F","#FFB000", "#FE6100")
pal3 = c("#DC267F","#FFB000", "#FE6100")
pal2 = c("#0947EA","#92EA33")
shape2 = c(21, 24)
plot_levels = c(
  "Control",
  "Thomas Fire Ash",
  "Low Temp. Ash",
  "High Temp. Ash",
  "Ash leachate",
  "DOC",
  "bold(Leachate)",
  "bold(Amendment)",
  "Low~biomass",
  "High~biomass",
  "Low biomass",
  "High biomass",
  "bold(Cells~(L^-1))",
  "bold(TOC~(µmol~C~L^-1))",
  "bold(BOC~(µmol~C~L^-1))",
  "bold(DOC~(µmol~C~L^-1))",
  "bold(Specific~growth~rate~(d^-1))",
  "bold(BGE~(dimensionless))",
  "bold(+DOC~bioav.~('%'))",
  "PB16",
  "PB15",
  "PB1",
  "PB14",
  "PB2",
  "PB3"
) 
```



```{r bacteria curves, fig.width=32, fig.asp=0.5}
curves <- ggplot(data4curves[!is.na(data4curves$val), ] %>% filter(var %in% c("bold(BOC~(µmol~C~L^-1))", "bold(DOC~(µmol~C~L^-1))")),
                 aes(
                   x = days,
                   y = val,
                   group = interaction(exp, trt, var),
                   color = factor(trt, levels = plot_levels),
                   fill = factor(trt, levels = plot_levels)
                 )) +
  geom_errorbar(
    aes(ymin = val - sd, ymax = val + sd),
    width = 0.2,
    linewidth = 0.6,
    alpha = 0.9
  ) +
  geom_line(linewidth = 1.1, alpha = 0.8) +
  geom_point( aes(shape = bge_inclusion),
    color = "black",
    size = 8,
    alpha = 0.8
  ) +
  scale_shape_manual(values = shape2, guide = guide_legend(override.aes = list(fill = "black"))) +
  scale_color_manual(values = pal4) +
  scale_fill_manual(values = pal4) +
  ggh4x::facet_nested(
    factor(var, levels = plot_levels) ~  factor(cluster, levels = plot_levels) + exp,
    scales = "free_y",
    labeller = label_parsed
  ) +
  labs(y = "", 
    x = expression(bold(Day)),
    fill = "",
    color = "",
    linetype = "",
    shape = "BGE calculation inclusion"
  ) +
  guides(color = "none", fill = guide_legend( 
    override.aes=list(shape = 21))) +
  theme_linedraw() +
  custom.theme 
```


```{r bact mu model, fig.width=26, fig.asp = 0.6}
# Fit the ANOVA model with interaction
mod1 <- aov(bact_mu ~ cluster * trt2, data = data4stats)

# Extract the ANOVA table
anova_results1 <- summary(mod1)

# Get p-values for significant terms
p_value_cluster <- anova_results1[[1]]["Pr(>F)"][1,1]
p_value_trt <- anova_results1[[1]]["Pr(>F)"][2,1]
# p_value_interaction <- anova_results2[[1]]["Pr(>F)"][3,1]

# Format p-values for annotations
formatted_p_cluster <- paste0("p = ", format.pval(p_value_cluster, digits = 2))
formatted_p_trt <- paste0("p = ", format.pval(p_value_trt, digits = 2))
# formatted_p_interaction <- paste0("p (cluster:trt) = ", format.pval(p_value_interaction, digits = 2))

cluster <- ggplot(data4stats, aes(x = factor(cluster, levels = plot_levels), y = bact_mu, fill = factor(cluster, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(cluster, levels = plot_levels)),
    size = 7,
    position = position_jitterdodge(),
    alpha = 0.5, 
    shape = 21,
    color = "black"
  ) +
  labs(
    x = "SOM cluster",
    y = expression(bold(Bacterial~specific~growth~rate~(d^-1))),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal2) +
  scale_color_manual(values = pal2) +
   theme_linedraw() +
  custom.theme +
  annotate(
    "text",
    x = 1,
    y = max(data4stats$bact_mu, na.rm = TRUE),
    label = formatted_p_cluster,
    size = 6,  vjust = 2
  ) +
   guides(color = "none") 


# Plot for Treatment Effect with ANOVA p-value using Box Plot
trt <-  ggplot(data4stats, aes(x = factor(trt2, levels = plot_levels), y = bact_mu, fill = factor(trt, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(trt, levels = plot_levels)),
    size = 7,
    shape = 21, 
    position = position_jitterdodge(),
    alpha = 0.5,
    color = "black"
  ) +
  labs(
    x = "Treatment",
    y = expression(bold(Bacterial~specific~growth~rate~(d^-1))),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal4) +
  scale_color_manual(values = pal4) +
   theme_linedraw() +
  custom.theme +
  # Add p-value annotations for interaction
  annotate(
    "text",
    x = 1,
    y = max(data4stats$bact_mu, na.rm = TRUE),
    label = formatted_p_trt,
    size = 6,  vjust = 2
  ) +
   guides(color= "none") 

# Combine plots side by side
mu <- cluster + trt +  plot_layout(guides = 'collect') & theme(legend.position = "top")

```
```{r bge model, fig.width=26, fig.asp = 0.6}
# Fit the ANOVA model with interaction
mod2 <- aov(bge ~ cluster * trt2, data = data4stats)

# Extract the ANOVA table
anova_results2 <- summary(mod2)

# Get p-values for significant terms
p_value_cluster <- anova_results2[[1]]["Pr(>F)"][1,1]
p_value_trt <- anova_results2[[1]]["Pr(>F)"][2,1]
# p_value_interaction <- anova_results2[[1]]["Pr(>F)"][3,1]

# Format p-values for annotations
formatted_p_cluster <- paste0("p = ", format.pval(p_value_cluster, digits = 2))
formatted_p_trt <- paste0("p = ", format.pval(p_value_trt, digits = 2))
# formatted_p_interaction <- paste0("p (cluster:trt) = ", format.pval(p_value_interaction, digits = 2))

cluster2 <- ggplot(data4stats, aes(x = factor(cluster, levels = plot_levels), y = bge, fill = factor(cluster, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(cluster, levels = plot_levels)),
    size = 7,
    position = position_jitterdodge(),
    alpha = 0.5, 
    shape = 21,
    color = "black"
  ) +
  labs(
    x = "SOM cluster",
    y = expression(bold(BGE)),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal2) +
  scale_color_manual(values = pal2) +
   theme_linedraw() +
  custom.theme +
  annotate(
    "text",
    x = 1,
    y = max(data4stats$bge, na.rm = TRUE),
    label = formatted_p_cluster,
    size = 6,  vjust = 2
  ) +
   guides(color = "none") 


# Plot for Treatment Effect with ANOVA p-value using Box Plot
trt2 <-  ggplot(data4stats, aes(x = factor(trt2, levels = plot_levels), y = bge, fill = factor(trt, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(trt, levels = plot_levels)),
    size = 7,
    shape = 21, 
    position = position_jitterdodge(),
    alpha = 0.5,
    color = "black"
  ) +
  labs(
    x = "Treatment",
    y = expression(bold(BGE)),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal4) +
  scale_color_manual(values = pal4) +
   theme_linedraw() +
  custom.theme +
  # Add p-value annotations for interaction
  annotate(
    "text",
    x = 1,
    y = max(data4stats$bge, na.rm = TRUE),
    label = formatted_p_trt,
    size = 6,  vjust = 2
  ) +
   guides(color= "none") 

# Combine plots side by side
bge <- cluster2 + trt2 +  plot_layout(guides = 'collect') & theme(legend.position = "top")

```

```{r Figure3, fig.width=37, fig.asp= 1.42, warning = F, message=F}
fig3 <- (
  (curves)
) /(mu)/(bge) + plot_layout(heights = c(0.8, 0.7, 0.7), guides = "collect") + plot_annotation(tag_levels = 'A')  &
  theme(plot.tag = element_text(size = 45), legend.position = "top", legend.box="vertical")

ggsave(fig3, file = "~/github/oceprf_smokeonthewater/prod/figs/bact.svg", width = 37, height = 48)
```



```{r doc rate, fig.width=26, fig.asp = 0.6}
# Fit the ANOVA model with interaction
mod3 <- aov(doc_rate ~ cluster * trt2, data = data4stats)

# Extract the ANOVA table
anova_results3 <- summary(mod3)

# Get p-values for significant terms
p_value_cluster <- anova_results3[[1]]["Pr(>F)"][1,1]
p_value_trt <- anova_results3[[1]]["Pr(>F)"][2,1]
p_value_interaction <- anova_results3[[1]]["Pr(>F)"][3,1]

# Format p-values for annotations
formatted_p_cluster <- paste0("p = ", format.pval(p_value_cluster, digits = 2))
formatted_p_trt <- paste0("p = ", format.pval(p_value_trt, digits = 2))
formatted_p_interaction <- paste0("p (Cluster : Treatment) = ", format.pval(p_value_interaction, digits = 2))

cluster <- ggplot(data4stats, aes(x = factor(cluster, levels = plot_levels), y = doc_rate, fill = factor(cluster, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(cluster, levels = plot_levels)),
    size = 7,
    position = position_jitterdodge(),
    alpha = 0.5, 
    shape = 21,
    color = "black"
  ) +
  labs(
    x = "SOM cluster",
    y = expression(bold(DOC~remineralization~rate~(µmol~C~L^-1~d^-1))),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal2) +
  scale_color_manual(values = pal2) +
   theme_linedraw() +
  custom.theme +
   guides(color = "none") 


# Plot for Treatment Effect with ANOVA p-value using Box Plot
trt <-  ggplot(data4stats, aes(x = factor(trt2, levels = plot_levels), y = doc_rate, fill = factor(trt, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(trt, levels = plot_levels)),
    size = 7,
    shape = 21, 
    position = position_jitterdodge(),
    alpha = 0.5,
    color = "black"
  ) +
  labs(
    x = "Treatment",
    y = expression(bold(DOC~remineralization~rate~(µmol~C~L^-1~d^-1))),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal4) +
  scale_color_manual(values = pal4) +
   theme_linedraw() +
  custom.theme +
  # Add p-value annotations for interaction
  annotate(
    "text",
    x = 1.5,  # Position for p-value of interaction
    y = max(data4stats$doc_rate, na.rm = TRUE) * 0.95,
    label = formatted_p_interaction,
    size = 6, vjust = 2
  ) +
  annotate(
    "text",
    x = 1.5,
    y = max(data4stats$doc_rate, na.rm = TRUE) * 0.85,
    label = formatted_p_cluster,
    size = 6, vjust = 2
  ) +
  annotate(
    "text",
    x = 1.5,
    y = max(data4stats$doc_rate, na.rm = TRUE) * 0.75,
    label = formatted_p_trt,
    size = 6, vjust = 2
  ) +
   guides(color= "none") 

# Combine plots side by side
doc_rate <- cluster + trt +  plot_layout(guides = 'collect') & theme(legend.position = "top")

```
```{r doc remin, fig.width=26, fig.asp = 0.6}
# Fit the ANOVA model with interaction
mod4 <- aov(amend_doc_bioav ~ cluster * trt, data = data4stats %>% filter(!trt == "Control"))

# Extract the ANOVA table
anova_results4 <- summary(mod4)

# Get p-values for significant terms
p_value_cluster <- anova_results4[[1]]["Pr(>F)"][1,1]
p_value_trt <- anova_results4[[1]]["Pr(>F)"][2,1]
# p_value_interaction <- anova_results4[[1]]["Pr(>F)"][3,1]

# Format p-values for annotations
formatted_p_cluster <- paste0("p = ", format.pval(p_value_cluster, digits = 2))
formatted_p_trt <- paste0("p = ", format.pval(p_value_trt, digits = 2))
# formatted_p_interaction <- paste0("p (Cluster : Treatment) = ", format.pval(p_value_interaction, digits = 2))

cluster <- ggplot(data4stats, aes(x = factor(cluster, levels = plot_levels), y = amend_doc_bioav, fill = factor(cluster, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(cluster, levels = plot_levels)),
    size = 7,
    position = position_jitterdodge(),
    alpha = 0.5, 
    shape = 21,
    color = "black"
  ) +
  labs(
    x = "SOM cluster",
    y = expression(bold(Bioavailability~of~amended~DOC~('%'))),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal2) +
  scale_color_manual(values = pal2) +
   theme_linedraw() +
  custom.theme +
   guides(color = "none") 


# Plot for Treatment Effect with ANOVA p-value using Box Plot
trt <-  ggplot(data4stats %>% filter(trt != "Control"), aes(x = factor(trt, levels = plot_levels), y = amend_doc_bioav, fill = factor(trt, levels = plot_levels))) +
  geom_boxplot(alpha = 0.7,
    outlier.shape = NA,
    width = 0.4,
    position = position_dodge(0.5)) +
  geom_jitter(
    aes(fill = factor(trt, levels = plot_levels)),
    size = 7,
    shape = 21, 
    position = position_jitterdodge(),
    alpha = 0.5,
    color = "black"
  ) +
  labs(
    x = "Treatment",
    y = expression(bold(Bioavailability~of~amended~DOC~('%'))),
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = pal3) +
  scale_color_manual(values = pal3) +
   theme_linedraw() +
  custom.theme +
  # Add p-value annotations for interaction
  annotate(
    "text",
    x = 1.5,  # Position for p-value of interaction
    y = max(data4stats$amend_doc_bioav, na.rm = TRUE) * 0.95,
    label = formatted_p_interaction,
    size = 6, vjust = 2
  ) +
  annotate(
    "text",
    x = 1.5,
    y = max(data4stats$amend_doc_bioav, na.rm = TRUE) * 0.85,
    label = formatted_p_cluster,
    size = 6, vjust = 2
  ) +
  annotate(
    "text",
    x = 1.5,
    y = max(data4stats$amend_doc_bioav, na.rm = TRUE) * 0.75,
    label = formatted_p_trt,
    size = 6, vjust = 2
  ) +
   guides(color= "none") 

# Combine plots side by side
bioav <- cluster + trt +  plot_layout(guides = 'collect') & theme(legend.position = "top")

```


```{r Figure4, fig.width=37, fig.asp= 1.42, warning = F, message=F}
fig4 <- (
 doc_rate)/(bioav) + plot_layout(heights = c(0.7, 0.7), guides = "collect") + plot_annotation(tag_levels = 'A')  &
  theme(plot.tag = element_text(size = 45), legend.position = "top", legend.box="vertical")

ggsave(fig4, file = "~/github/oceprf_smokeonthewater/prod/figs/doc.svg", width = 37, height = 40)
```


# Supplementary Table

```{r Supp Table1}
table_data <- p5_prod %>% 
  select(exp, trt, days, interp_bocdoc, bge_inclusion, interp_fgC_cell, boc_rate, doc_rate) %>% 
  mutate_at(vars(contains(c("rate"))), round, 2) %>% 
  mutate_at(vars(contains("interp_fgC_cell")), round, 1) %>% 
  mutate(bge_inclusion = ifelse(is.na(bge_inclusion), F, bge_inclusion))
  
gt_tbl <- gt(table_data) 

table <- 
  gt_tbl |>
  tab_header(
    title = md("**Table S2.** Cell-specific bacterial carbon, bacterial organic carbon rate changes, and DOC removal rates"),
  ) |>
  cols_label(
    exp = html("Experimental site"),
    trt = html("Treatment"),
    days = html("Days"), 
    interp_bocdoc = html("Interpolated cell-specific bacterial carbon"),
    bge_inclusion = html("BGE calculation inclusion"),
    interp_fgC_cell = html("Cell-specific bacterial carbon"), 
    boc_rate = html("Rate change in bacterial organic carbon"),
    doc_rate = html("Rate change in DOC"),
  ) |>
  tab_spanner(
    label = html("µmol C L<sup>-1</sup> d<sup>-1"),
    columns = c(boc_rate, doc_rate)
  ) |>
  tab_spanner(
    label = html("fg C cell<sup>-1"),
    columns = c(interp_fgC_cell)
  ) 

table_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/knitted/5_bacteria_dom_exps_files/bact_table.html"

gtsave(table, table_path)
```

# Supplementary Figure

```{r supplemental curves, fig.width=32, fig.asp=0.5}
suppfig1 <- ggplot(data4curves[!is.na(data4curves$val), ] %>% filter(!var %in% c("bold(BOC~(µmol~C~L^-1))", "bold(DOC~(µmol~C~L^-1))")),
                 aes(
                   x = days,
                   y = val,
                   group = interaction(exp, trt, var),
                   color = factor(trt, levels = plot_levels),
                   fill = factor(trt, levels = plot_levels)
                 )) +
  geom_errorbar(
    aes(ymin = val - sd, ymax = val + sd),
    width = 0.2,
    linewidth = 0.6,
    alpha = 0.9
  ) +
  geom_line(linewidth = 1.1, alpha = 0.8) +
  geom_point( aes(shape = bge_inclusion),
    color = "black",
    size = 8,
    alpha = 0.8
  ) +
  scale_shape_manual(values = shape2, guide = guide_legend(override.aes = list(fill = "black"))) +
  scale_color_manual(values = pal4) +
  scale_fill_manual(values = pal4) +
  ggh4x::facet_nested(
    factor(var, levels = plot_levels) ~  factor(cluster, levels = plot_levels) + exp,
    scales = "free_y",
    labeller = label_parsed
  ) +
  labs(y = "", 
    x = expression(bold(Day)),
    fill = "",
    color = "",
    linetype = "",
    shape = "BGE calculation inclusion"
  ) +
  guides(color = "none", fill = guide_legend( 
    override.aes=list(shape = 21))) +
  theme_linedraw() + 
  custom.theme 

ggsave(suppfig1, file = "~/github/oceprf_smokeonthewater/prod/figs/bact_suppl.svg", width = 37, height = 20)
```

