---
title: "Bacteria and DOM experiments"
author: "Nick Baetge"
date: "compiled most recently on `r format(Sys.time(), '%d %B, %Y')`"
output: github_document

knit: (function(inputFile, encoding) {
      out_dir <- "../knitted";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
---

```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(hms)
library(lubridate)
library(purrr)
library(patchwork)
library(janitor)
```

```{r define path to save data}
prod_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/p5_bacteria_dom_exps.csv"
```

# Bacterial cell data

## Import and tidy data

```{r define paths to fcm data}
fcm_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/p4_fcm_plate_data"
filter_vols <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_filter_vols.xlsx"
time_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_bact_dom_sampletimes.csv"
```

```{r import and tidy sample time data, message = F}
dt <- read_csv(time_path) %>%
  mutate_at(vars(datetime_utc), ymd_hms) %>%
  rename(dt = 3) %>%
  group_by(exp) %>%
  mutate(
    interv = interval(first(dt), dt),
    dur = as.duration(interv),
    days = round(time_length(dur, unit = "day"), 2)
  ) %>%
  select(exp, tp, dt, days)
```

import data and clean based on run notes:

- B2 run 1 (and dil): timepoint 1 and 2 are switched
- B3_14 run 1 dil: B14-2-4-0 didn't add sample so data is in well H12
- B16 run 2 dil:  B16-V2-4-4 didn't add sample so data is in well G1
- B3_14 run 2 B14_V1_3 was a 2 ml sample instead of a 1 ml sample, also extra 1.2 vial (could be 1.2 GF75)
- Filtrates run 2: missing vial for B3 1.2 GF75 

```{r import and tidy fcm data, message = F}
fcm <-
  list.files(path = fcm_path,
             pattern = "*.csv",
             full.names = T) %>%
  map_df( ~ read_csv(., col_types = cols(.default = "c"))) %>%
  mutate_at(vars(tp, dil, cells_ml), as.numeric) %>%
  mutate(tp = ifelse(
    name %in% c(
      "B2_dil-1.fcs",
      "B2-7.fcs",
      "B2_dil-2.fcs",
      "B2-8.fcs",
      "B2_dil-3.fcs",
      "B2-9.fcs",
      "B2_dil-4.fcs",
      "B2-10.fcs"
    ),
    2,
    tp
  ),
  tp = ifelse(
    name %in% c("B2-11.fcs", "B2-12.fcs", "B2-13.fcs", "B2-14.fcs"),
    1,
    tp
  )) %>%   # sample switch
  mutate(btl = ifelse(name == "B3_14-1.fcs", "1.2GF75", btl)) %>%    # sample switch
  filter(!name %in% c(
    "B3_14_dil-33.fcs",
    "B15-2.fcs",
    "B16_r2-25.fcs",
    "B16_r2-25.fcs"
  )) %>%    # samples had 0 events
  mutate(
    trt = case_when(
      btl %in% c("1", "V1", "1.2T", "1.2DNA", "1.2GF75", "1DNA", "1GF75") ~ 1,
      btl %in% c("2", "V2", "2DNA", "2GF75") ~ 2,
      btl %in% c("3", "V3", "3DNA", "3GF75") ~ 3,
      btl %in% c("4", "V4", "4DNA", "4GF75") ~ 4
    ),
    type = ifelse(btl %in% c("1", "2", "3", "4"), "5L Bottle", "40 mL Vial"),
    desc = case_when(
      trt == 1 ~ "Control",
      trt == 2 ~ "Thomas Fire Ash",
      trt == 3 ~ "Low Temp. Ash",
      trt == 4 ~ "High Temp. Ash"
    ),
    .after = btl
  ) %>%
  arrange(exp, trt, tp) 
```

## Subtract sybr stain

```{r calculate sybr correction}
sybr <- fcm %>%
  filter(exp == "TE+SYBR") %>%
  mutate(med_sybr = round(median(cells_ml))) %>%
  select(med_sybr) %>%
  distinct() %>%
  as.numeric()
```

```{r correct for sybr}
fcm_corr <- fcm %>%
  select(name, exp, btl, trt, desc, type, tp, dil, cells_ml) %>%
  arrange(exp, btl, tp) %>%
  filter(!exp %in% c("TE", "TE+SYBR")) %>%
  mutate(cells_ml = cells_ml - sybr,
         cells_ml = ifelse(cells_ml < 0, 0, cells_ml)) 
```

## Assess vial v bottle cell dynamics

```{r regression of vial cells and bottle cells, message = F}
reg_data <- fcm %>%
  filter(btl %in% c("1", "2", "3", "4", "V1", "V2", "V3", "V4")) %>%
  group_by(exp, trt, btl, type, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml))) %>%
  ungroup() %>%
  select(exp, trt, desc, type, tp, dil, mean_cells_ml) %>%
  distinct() %>%
  pivot_wider(
    id_cols = c(exp, trt, desc, tp, dil),
    names_from = type,
    values_from = mean_cells_ml
  ) %>%
  drop_na(7)

reg <-
  lmodel2::lmodel2(reg_data$`40 mL Vial` ~ reg_data$`5L Bottle`, nperm = 99)

reg
```
The cell dynamics occurring in the bottle and the vials track well with each other. we can use averages between the two for subsequent measurements. so cells per ml will reflect the average of two subsamples from the vials and two subsamples from the bottle (n = 4)

## Extract experimental data

```{r extract experimental data, message = F}
# apply cell counts from bottles at T0 as vial values for T0
fcm_40mlt0 <- fcm %>%
  filter(btl %in% c("1", "2", "3", "4")) %>%
  group_by(exp, trt, btl, type, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml)),
         sd_cells_ml = round(sd(cells_ml))) %>%
  ungroup() %>%
  filter(tp == 0) %>%
  mutate(type = "40 mL Vial",
         btl = paste("V", btl, sep = ""))

#extract experimental data and merge with datetime data
fcm_exp <- fcm %>%
  filter(btl %in% c("1", "2", "3", "4", "V1", "V2", "V3", "V4")) %>%
  group_by(exp, trt, btl, type, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml)),
         sd_cells_ml = round(sd(cells_ml))) %>%
  ungroup() %>%
  bind_rows(., fcm_40mlt0) %>%
  select(-name) %>%
  arrange(exp, btl, tp) %>%
  left_join(., dt) %>%
  select(exp:tp, dt, days, everything()) %>%
  mutate(hours = days * 24, .after = days)

#calculate mean data
fcm_exp_means <- fcm_exp %>%
  select(-c(type, dil, btl, mean_cells_ml, sd_cells_ml)) %>%
  arrange(exp, trt, tp) %>%
  group_by(exp, trt, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml)),
         sd_cells_ml = round(sd(cells_ml))) %>%
  select(-cells_ml) %>%
  distinct() %>% 
  ungroup() 
```

## Estimate growth rates (d-1) in exponential phase

```{r estimate growth rates, message = F}
fcm_gr <- fcm_exp %>%
  filter(
    exp == "B1" & tp <= 3 |
      exp == "B2" & tp <= 2 |
      exp == "B3" & tp <= 2 |
      exp == "B14" & tp <= 2 |
      exp == "B15" &  tp %in% c(1, 2, 3) |
      exp == "B16" & tp %in% c(1, 2, 3)
  ) %>%
  group_by(exp, btl, trt) %>%
  mutate(gr_d = (log(last(mean_cells_ml)) - log(first(mean_cells_ml))) / (last(days) - first(days))) %>%
  select(exp:desc, gr_d) %>%
  distinct() %>%
  group_by(exp, trt, desc) %>%
  summarise_at(vars(gr_d), list(mean = mean, sd = sd)) %>%
  ungroup() %>%
  rename(mean_gr_d = mean,
         sd_gr_d = sd)

p5_fcm <- fcm_gr  %>%
  left_join(fcm_exp_means, .) %>%
  select(exp,
         desc,
         tp,
         dt,
         days,
         hours,
         mean_cells_ml,
         sd_cells_ml,
         mean_gr_d,
         sd_gr_d) %>%
  distinct()
```

## Estimate filter retention for bacterial organic carbon and dna samples

```{r extract filtration data and calculate retention, message = F}
fcm_gf75_dna <- fcm %>%
  filter(tp %in% c(0, 2, 5)) %>%
  mutate(cells_ml = ifelse(
    name %in% c(
      "Filtrates-21.fcs",
      "Filtrates_r2-21.fcs",
      "Filtrates_r2-19.fcs",
      "Filtrates-19.fcs"
    ),
    NA,
    cells_ml
  )) %>%  # suspect samples. these gf75 filtrates had more cells than any other sample in the dataset.
  arrange(exp, tp, btl) %>%
  group_by(exp, btl, tp) %>%
  mutate(mean_cells_ml = round(mean(cells_ml))) %>%
  ungroup() %>%
  select(-c(name, dil, cells_ml)) %>%
  distinct() %>%
  filter(!btl %in% c(1, 2, 3, 4) | !tp == 0) %>%
  group_by(exp, trt, tp) %>%
  mutate(
    total = ifelse(
      str_detect(btl, "DNA") |
        str_detect(btl, "GF75"),
      NA,
      mean_cells_ml
    ),
    gf75_filtrate = ifelse(str_detect(btl, "GF75"), mean_cells_ml, NA),
    dna_filtrate = ifelse(str_detect(btl, "DNA"), mean_cells_ml, NA)
  ) %>%
  fill(c(total:dna_filtrate), .direction = "updown") %>%
  ungroup() %>%
  select(exp, trt, desc, tp, total:dna_filtrate) %>%
  distinct() %>%
  group_by(exp, trt, tp) %>%
  mutate(mean_total = round(mean(total, na.rm = T)), .after = tp) %>%
  select(-total) %>%
  distinct() %>%
  rename(total = mean_total) %>%
  left_join(readxl::read_xlsx(filter_vols, sheet = 1), .) %>%
  mutate(
    gf75_total = total * gf75_vol,
    dna_total = total * dna_vol,
    .before = gf75_filtrate
  ) %>% #total cells in 1L
  mutate_at(vars(dna_filtrate), ~ . * dna_vol) %>%
  mutate_at(vars(gf75_filtrate), ~ . * gf75_vol) %>%
  select(-total) %>%
  mutate(
    gf75_filter = gf75_total - gf75_filtrate,
    dna_filter = dna_total - dna_filtrate,
    gf75_percent_retention = round(gf75_filter / gf75_total * 100),
    dna_percent_retention = round(dna_filter / dna_total * 100)
  )
```

```{r filter retention of cells summary statistics}
summary(fcm_gf75_dna)
sd(fcm_gf75_dna$gf75_percent_retention, na.rm = T)
sd(fcm_gf75_dna$dna_percent_retention)
```

# Bacterial organic carbon data

## Import and tidy data

```{r define paths to boc data}
chn_path <- "/Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_chn.csv"
```

```{r import data, message = F, warning = F}
p5_boc <- read_csv(chn_path) %>% 
  left_join(., readxl::read_xlsx(filter_vols, sheet = 2)) %>% 
  select(1, 2, 4, 5, 7) %>% 
  rename(rundate = 1,
         exp_sample_tp_rep = 2, 
         c_ug = 3, 
         n_ug = 4) %>% 
  separate(., 2, into = c("exp", "sample", "tp", "rep"), sep = "-") %>% 
  separate(., 3, into = c("sample", "filter"), sep = -1) %>% 
  group_by(exp, sample, tp, rep) %>% 
   mutate(raw_c = round(sum(c_ug),1),
         raw_n = round(sum(n_ug),1)) %>% 
  select(-c(filter, c_ug, n_ug)) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(c_blank = round(mean(raw_c[exp == "BLK"], na.rm = T), 1),
         n_blank = round(mean(raw_n[exp == "BLK"], na.rm = T), 1),
         c_blank_sd = round(sd(raw_c[exp == "BLK"], na.rm = T), 1),
         n_blank_sd = round(sd(raw_n[exp == "BLK"], na.rm = T), 1)) %>% 
  filter(!exp == "BLK") %>% 
  mutate(c_ug = raw_c - c_blank,
         n_ug = raw_n - n_blank) %>% 
  filter(!c_ug < 0) %>% 
  mutate(c_mg = c_ug / 10^3,
         n_mg = n_ug / 10^3,
         c_umol = c_mg * 10^3 / 12,
         n_umol = n_mg * 10^3 / 14,
         vol_m3 = vol / 10^6,
         c_mg_m3 = round(c_mg / vol_m3, 1),
         n_mg_m3 = round(n_mg / vol_m3, 1),
         c_umol_l = c_mg_m3/12000,
         n_umol_l = n_mg_m3/14000,
         cn =  round(c_umol_l/n_umol_l, 1)) %>% 
  select(exp, sample, tp, c_mg:cn) %>% 
  rename(trt = sample) %>% 
  group_by(exp, trt, tp) %>% 
  summarize(across(
    .cols = c_mg:cn,
    .fns = list(mean = mean, sd = sd),
    na.rm = TRUE,
    .names = "{fn}_{col}"
  )) %>% 
  ungroup() %>% 
  mutate_at(vars(tp), as.numeric) %>% 
  mutate(trt = ifelse(trt == 1.2, 1, trt)) %>% 
  group_by(exp, trt) %>% 
  group_modify(~ add_row(., tp = 0 )) %>% 
  arrange(exp,  tp) %>% 
  group_by(exp, tp) %>% 
  fill(c(3:23), .direction = "down") %>% 
  ungroup() %>% 
  distinct() %>% 
  arrange(exp, trt, tp) %>% 
  mutate_at(vars(trt), as.numeric) %>% 
  left_join(., fcm_gf75_dna %>% select(exp, trt, desc, tp, gf75_filter)) %>% 
  group_by(exp, tp) %>% 
  fill(gf75_filter, .direction = "downup") %>% 
  ungroup() %>% 
  rename(gf75_cells = gf75_filter) %>% 
  select(exp, trt, desc, tp, gf75_cells, everything()) %>% 
  mutate(fgC_cell = (mean_c_umol * 12 * 10^9)/gf75_cells,
         fgN_cell = (mean_n_umol * 14 * 10^9)/gf75_cells ) %>% 
  mutate(desc = case_when(trt == 1 ~ "Control",
                          trt == 2 ~ "Thomas Fire Ash",
                          trt == 3 ~ "Low Temp. Ash",
                          trt == 4 ~ "High Temp. Ash")) %>% 
  select(exp, trt, desc, tp, fgC_cell, fgN_cell) %>% 
  left_join(p5_fcm,.) %>%
  mutate(boc = (mean_cells_ml * fgC_cell) / (12 * 10^6),
         bon = (mean_cells_ml * fgN_cell) / (14 * 10^6)) %>% 
  select(-trt)  %>% 
  rename(trt = desc)
```

c_blank = 15.3 +/- 5.1, n = 9
n_blank = 7.1 +/- 1, n = 9 

# DOC data

## Import and tidy data

```{r define paths to doc data}
doc_path <- "//Users/nicholasbaetge/github/oceprf_smokeonthewater/raw/r5_doc.xlsx"
```

```{r tidy and derive bge and doc bioav, message = F}
doc <- readxl::read_xlsx(doc_path, skip = 27) %>%
  drop_na() %>%
  select(1, 3) %>%
  rename(id = 1,
         toc = 2) %>%
  separate(id, into = c("exp", "btl", "tp"), sep = "-") %>%
  group_by(exp, btl, tp) %>%
  mutate(mean_toc = mean(toc),
         sd_toc = sd(toc),
         tp = as.numeric(tp)) %>%
  ungroup() %>%
  select(-toc) %>%
  distinct() %>%
  mutate(
    trt =  case_when(
      btl == 1 ~ "Control",
      btl == 2 ~ "Thomas Fire Ash",
      btl == 3 ~ "Low Temp. Ash",
      btl == 4 ~ "High Temp. Ash"
    ),
    .after = btl
  ) %>%
  select(-btl) %>%
  left_join(p5_boc, .) %>%
  mutate(doc = mean_toc - boc)

#calculate boc rate of change as slope
boc_rate <- doc %>% 
  select(exp, trt, tp, days, boc) %>%
  drop_na(boc) %>%
  group_by(exp, trt) %>%
  mutate(group_id = cur_group_id()) %>%
  ungroup() %>%
  group_split(group_id) %>%
  map( ~ lm(boc ~ days, data = .x)) %>%
  map_df(broom::tidy, .id = "group_id") %>%
  filter(term == "days") %>%
  select(group_id, estimate) %>%
  rename(boc_rate = estimate) %>%
  mutate_at(vars(group_id), as.numeric) %>%
  ungroup() %>% 
  left_join(
    .,
    doc %>%
      select(exp, trt) %>% distinct() %>%
      group_by(exp, trt) %>%
      mutate(group_id = cur_group_id()) %>%
      ungroup()
  ) %>% 
  select(-group_id)

#calculate doc rate of change as slope
doc_rate <- doc %>%
  select(exp, trt, tp, days, doc) %>%
  drop_na(doc) %>%
  group_by(exp, trt) %>%
  mutate(group_id = cur_group_id()) %>%
  ungroup() %>%
  group_split(group_id) %>%
  map( ~ lm(doc ~ days, data = .x)) %>%
  map_df(broom::tidy, .id = "group_id") %>%
  filter(term == "days") %>%
  select(group_id, estimate) %>%
  rename(doc_rate = estimate) %>%
  mutate_at(vars(group_id), as.numeric) %>%
  mutate_at(vars(doc_rate), abs) %>%
  ungroup() %>% 
  left_join(
    .,
    doc %>%
      select(exp, trt) %>% distinct() %>%
      group_by(exp, trt) %>%
      mutate(group_id = cur_group_id()) %>%
      ungroup()
  ) %>% 
  select(-group_id)

#calculate bges and doc bioavailability
p5_bge_bioav <- doc %>% 
  left_join(., boc_rate) %>% 
  left_join(., doc_rate) %>% 
  group_by(exp, trt) %>% 
  mutate(doc_rem = ifelse(is.na(doc[tp == 5]), first(doc) - doc[tp == 2], first(doc) - last(doc) ),
         doc_bdl = ifelse(doc_rem <= 1.4, T, F),
         bge = ifelse(doc_bdl == F, boc_rate/doc_rate, NA)) %>% 
  ungroup() %>% 
  group_by(exp, tp) %>% 
  mutate(control_doc = doc[trt == "Control"]) %>% 
  ungroup() %>% 
  mutate(amend_doc = doc - control_doc) %>% 
  group_by(exp, trt) %>% 
  mutate(amend_doc_bioav = ifelse(trt != "Control", (amend_doc[tp == 0] - amend_doc[tp == 5]) / amend_doc[tp == 0]* 100, NA )) %>% 
  ungroup() %>% 
  mutate(exp = paste("P", exp, sep = ""))
```

# Merge data with contextual in situ data 

```{r define paths to context data, message = F}
context_path <-
  "/Users/nicholasbaetge/github/oceprf_smokeonthewater/prod/p1_insitu_index.csv"

p5_prod <- read_csv(context_path) %>% 
  select(stn, exp, lat, lon, biomass, composite_z) %>% 
  left_join(., p5_bge_bioav) %>% 
  arrange(biomass, composite_z) %>% 
  drop_na(trt) %>% 
  mutate_at(vars(composite_z), round, 2) %>% 
  arrange(desc(composite_z), .by_group = TRUE) 
levels(p5_prod$composite_z)
```

# Save data

```{r save data}
write_csv(p5_prod, prod_path)
```

# Plot data

```{r custom theme for plots}
custom.theme <- theme(
  legend.position = "top",
  legend.title = element_text(size = 35),
  legend.key.size = unit(3, "cm"),
  legend.key.spacing.x = unit(0.75, "cm"),
  legend.text = element_text(size = 38),
  legend.box = "vertical",
  legend.margin = margin(),
  axis.title = element_text(size = 35, face = "bold"),
  axis.text = element_text(size = 28),
  panel.spacing.x = unit(0.5, "cm"),
  panel.spacing.y = unit(1, "cm"),
  strip.text.x = element_text(
    size = 35,
    color = "white",
    face = "bold"
  ),
  strip.text.y = element_text(
    size = 35,
    color = "white",
    face = "bold"
  ),
  strip.background = element_rect(
    color = "black",
    fill = alpha('black', 0.7),
    linewidth = 1.5,
    linetype = "solid"
  )
)

pal4 = c("#2C1B21", "#6aa42c", "#5990e7", "#b95f83")
pal3 = c("#6aa42c", "#5990e7", "#b95f83")
plot_levels = c(
  "Control",
  "Thomas Fire Ash",
  "Low Temp. Ash",
  "High Temp. Ash",
  "DOC",
  "bold(Leachate)",
  "bold(Amendment)",
  "bold(Lower~biomass~index)",
  "bold(Higher~biomass~index)",
  "bold(TOC~(µmol~C~L^-1))",
  "bold(Bacterial~C~(µmol~C~L^-1))",
  "bold(Cells~(L^-1))",
  "bold(Growth~rate~(d^-1))",
  "bold(BGE~(dimensionless))",
  "bold(+DOC~bioav.~('%'))",
  "PB16",
  "PB15",
  "PB1",
  "PB14",
  "PB2",
  "PB3"
) 
```

```{r pivot data for figure 4, message = F}
fig4_data <- p5_prod %>%
  select(exp, trt, biomass, composite_z, days, mean_toc, boc) %>%
  rename(toc = mean_toc) %>%
  pivot_longer(toc:boc, names_to = "var", values_to = "val") %>%
  mutate(
    var = case_when(
      var == "toc" ~ "bold(TOC~(µmol~C~L^-1))",
      var == "boc" ~ "bold(Bacterial~C~(µmol~C~L^-1))"
    )
  ) %>%
  left_join(
    .,
    p5_prod %>%
      select(exp, trt, biomass, days,  composite_z, sd_toc) %>%
      mutate(var = "bold(TOC~(µmol~C~L^-1))") %>%
      rename(sd = sd_toc)
  ) %>%
  bind_rows(
    .,
    p5_prod %>%
      select(exp, trt, biomass, days,  composite_z, mean_cells_ml, sd_cells_ml) %>%
      mutate_at(vars(mean_cells_ml, sd_cells_ml), ~ . * 10 ^ 3) %>%
      mutate(var = "bold(Cells~(L^-1))") %>%
      rename(val = mean_cells_ml,
             sd = sd_cells_ml)
  ) %>%
  arrange(exp, trt, days, var) %>%
  mutate(
    biomass = ifelse(
      biomass == "Lower",
      "bold(Lower~biomass~index)",
      "bold(Higher~biomass~index)"
    )
  ) 
```

```{r bacteria oc curves, fig.width=32, fig.asp=0.7}
curves <- ggplot(fig4_data[!is.na(fig4_data$val), ],
                 aes(
                   x = days,
                   y = val,
                   group = interaction(exp, trt, var),
                   color = factor(trt, levels = plot_levels),
                   fill = factor(trt, levels = plot_levels)
                 )) +
  geom_errorbar(
    aes(ymin = val - sd, ymax = val + sd),
    width = 0.2,
    linewidth = 0.6,
    alpha = 0.9
  ) +
  geom_line(linewidth = 1.1, alpha = 0.8) +
  geom_point(
    color = "black",
    size = 8,
    shape = 21,
    alpha = 0.8
  ) +
  scale_color_manual(values = pal4) +
  scale_fill_manual(values = pal4) +
  ggh4x::facet_nested(
    factor(var, levels = plot_levels) ~  factor(biomass, levels = plot_levels) + factor(
      composite_z,
       levels = c("-5.9", "-3.48", "-0.77", "0.09", "0.94", "0.96")
    ),
    scales = "free_y",
    labeller = label_parsed
  ) +
  labs(
    y = "",
    x = expression(Day),
    fill = "",
    color = "",
    linetype = ""
  ) +
  guides(color = "none") +
  theme_linedraw() +
  custom.theme 
```

```{r pivot data for derived metrics, message = F}
metric_data <- p5_prod %>%
  select(exp,
         trt,
         biomass,
         composite_z,
         mean_gr_d,
         bge,
         amend_doc_bioav) %>%
  distinct() %>%
  pivot_longer(mean_gr_d:amend_doc_bioav,
               names_to = "var",
               values_to = "val") %>%
  mutate(
    var = case_when(
      var == "mean_gr_d" ~ "bold(Growth~rate~(d^-1))",
      var == "bge" ~ "bold(BGE~(dimensionless))",
      var == "amend_doc_bioav" ~ "bold(+DOC~bioav.~('%'))",
    )
  ) %>%
  left_join(
    .,
    p5_prod %>%
      select(exp, trt, biomass,  composite_z, sd_gr_d) %>%
      distinct() %>%
      mutate(var = "bold(Growth~rate~(d^-1))") %>%
      rename(sd = sd_gr_d)
  ) %>%
  mutate(biomass = ifelse(
    biomass == "Lower",
    "bold(Lower~biomass~index)",
    "bold(Higher~biomass~index)"
  ))
```

```{r bact oc metrics, fig.width=26, fig.asp = 0.8, warning = F, message = F}
metrics <- ggplot(metric_data,
                  aes(
                    x = factor(
                      composite_z,
                      levels = c("-5.9", "-3.48", "-0.77", "0.09", "0.94", "0.96")
                    ),
                    y = val,
                    fill = factor(trt, levels = plot_levels),
                    color = factor(trt, levels = plot_levels)
                  )) +
  geom_errorbar(aes(ymin = val - sd,
                    ymax = val + sd),
                width = .1,
                position = position_dodge(.4)) +
  geom_linerange(
    aes(ymin = 0, ymax = val),
    position = position_dodge(width = 0.4),
    linewidth = 1,
    alpha = 0.8
  ) +
  geom_point(position = position_dodge(width = 0.4),
             size = 10,
             alpha = 0.8) +
  ggh4x::facet_nested_wrap(
    ~ factor(var, levels = plot_levels),
    labeller = label_parsed,
    scales = "free",
    strip.position = "right",
    nrow = 3
  ) +
  scale_fill_manual(values = pal4) +
  scale_color_manual(values = pal4) +
  labs(y = expression(""),
       x = expression("Biomass index"),
       color = "") +
  theme_linedraw() +
  custom.theme +
  guides(fill = "none") +
  theme(panel.spacing.y = unit(0.5, "cm"))
```

```{r Figure3, fig.width=32, fig.asp= 1.42, warning = F, message=F}
(
  curves
) /(metrics  + guides(color = "none", fill = "none")) + plot_layout(heights = c(0.8, 0.7)) + plot_annotation(tag_levels = 'A')  &
  theme(plot.tag = element_text(size = 28))
```
